---
title: "ISYE 6501 HWK 2"
editor_options:
  markdown:
    wrap: 100
output: html_document
#date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Import Packages and IDE Prep**
```{r}
#rm(list = ls())
set.seed(2)
library(kernlab, plyr)
library(kknn)
library(tidyverse)
library(stats)
options(scipen=999)
```

**Read in Dataset(s)**
```{r}
data(iris)
head(iris,10)
```
**Exploratory Data Analysis**
```{r}
#dimensions
dim = dim(iris)
dim

#View unique values in response variable (species)
unique_species = unique(iris['Species'])
unique_species

#summary stats on IRIS
summary(iris)
```

**Visualize data**
```{r}
ggplot(data = iris, aes(Sepal.Length, Sepal.Width, color = Species)) + geom_point()
ggplot(data = iris, aes(Petal.Length, Petal.Width, color = Species)) + geom_point()
```


**Scale IRIS dataset**
```{r}
##scale data
iris_scaled = scale(iris[,1:4], center = TRUE, scale = TRUE)
iris_scaled = as.data.frame(iris_scaled)
iris_scaled['Species'] = iris[,5]
head(iris_scaled,5)

```
**Visualize Scaled IRIS DS --Note there is clearer clustering with Petal Length & Petal Width**
```{r}
ggplot(data = iris_scaled, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point()
ggplot(data = iris_scaled, aes(Petal.Length, Petal.Width, color = Species)) +
  geom_point()

```

**In order to find the optimal number of centers lets make an elbow diagram using All 4 predictors**
```{r}
#for loop to grab the total within sum of squares from each kmeans model with varying number of centers
results_df_kmeans_elbow = data.frame(center = numeric(), totwithinss = numeric()) 
for (center in seq(1,10)) {
  kmeans = kmeans(iris_scaled[,1:4], centers = center, nstart = 10)
  totwithinss = kmeans$tot.withinss
  results_df_kmeans_elbow = rbind(results_df_kmeans_elbow, data.frame(center = center, totwithinss = totwithinss)) 
}

plot(x = results_df_kmeans_elbow$center, y = results_df_kmeans_elbow$totwithinss
     ,type = 'b', xlab = 'Number of Clusters', ylab = 'Total_Within_Sum_Squares')

```

**Elbow Diagram using only Petal predictors based on earlier graphs where there is better clustering using only petal predictors**
```{r}
#for loop to grab the total within sum of squares from each kmeans model with varying number of centers
results_df_kmeans_elbow_petals = data.frame(center = numeric(), totwithinss = numeric()) 
for (center in seq(1,10)) {
  kmeans = kmeans(iris_scaled[,3:4], centers = center, nstart = 10)
  totwithinss = kmeans$tot.withinss
  results_df_kmeans_elbow_petals = rbind(results_df_kmeans_elbow_petals, data.frame(center = center, totwithinss = totwithinss)) 
}


plot(x = results_df_kmeans_elbow_petals$center, y = results_df_kmeans_elbow_petals$totwithinss
     ,type = 'b', xlab = 'Number of Clusters', ylab = 'Total_Within_Sum_Squares')
```

**3 clusters seems to be the most promising based on the elbow diagram on the Petal predictors (petal.length & petal.width)**


**Lets view model for 3 clusters using only petal width and petal length as our predictors**
```{r}
kmeans_petalwl_3 = kmeans(iris[,3:4], centers = 3, nstart = 10)
kmeans_petalwl_3
```

**Graphing predictions (k = 3, Petal.Length & Petal.Width)**
```{r}
ggplot(iris, aes(Petal.Length, Petal.Width, color = kmeans_petalwl_3$cluster)) + geom_point()
```